---
title: "Lab 8"
output: html_document
date: "2023-11-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





```{r}
# for now, really simplify the education dummy
#adding college and advanced degree for people above a bachelors

#includes everyone, male and female who goes to college

data_use2$BA_plus <- data_use2$educ_college + data_use2$educ_advdeg


#whole data set

model_lpm_v1 <- lm(public_work_num ~ female + BA_plus + AGE + I(female*BA_plus) + I(AGE*female), data = data_use2)
summary(model_lpm_v1)

#turning the subsets for females and males into a logical argument

data_use2_female <- subset(data_use2,as.logical(data_use2$female))
data_use2_male <- subset(data_use2,!(data_use2$female))

# now split into 2 parts
model_lpm_v1f <- lm(public_work_num ~ BA_plus + AGE, data = data_use2_female)
summary(model_lpm_v1f)
model_lpm_v1m <- lm(public_work_num ~ BA_plus + AGE, data = data_use2_male)
summary(model_lpm_v1m)

```
# Is it the same between the interaction terms? 
explanatory variable is public work
Compare the results of overall data set(with multiplications) vs the results of just the female specific model to our model
The results of the coefficients were similar but the residuals were quite differen, the subset of only female was much smaller than the whole data set.

#
interpreting dummy variables and interactions
when x variable goes up by 1, y goes up by 
pluggig in a simple dummy so implicitly comparing with others

# use full data set 

```{r}
model_lpm_v2f <- lm(public_work_num ~ educ_college + educ_advdeg + AGE, data = data_use2_female)
summary(model_lpm_v2f)
model_lpm_v2m <- lm(public_work_num ~ educ_college + educ_advdeg + AGE, data = data_use2_male)
summary(model_lpm_v2m)
```


#
forest
tree setting up a number of decisions to classify creating branches

```{r}
require('randomForest')
set.seed(54321)

model_randFor <- randomForest(as.factor(pub_work) ~ ., data = sobj$data, importance=TRUE, proximity=TRUE)
print(model_randFor)

#confusion matrix

#how well it tests on data
round(importance(model_randFor),2)
varImpPlot(model_randFor)
# look at confusion matrix for this too

pred_model1 <- predict(model_randFor,  s_dat_test)
table(pred = pred_model1, true = dat_test$pub_work)
```



``` {r}
require(e1071)
# tuned_parameters <- tune.svm(as.factor(pub_work) ~ ., data = sobj$data, gamma = 10^(-3:0), cost = 10^(-2:2)) 
# summary(tuned_parameters)
# figure best parameters and input into next
svm.model <- svm(as.factor(pub_work) ~ ., data = sobj$data, cost = 1, gamma = 0.1)
svm.pred <- predict(svm.model, s_dat_test)
table(pred = svm.pred, true = dat_test$pub_work)
```


When you summarize, you should be able to explain which models predict best (noting if there is a tradeoff of false positive vs false negative) and if there are certain explanatory variables that are consistently more or less useful. Also try other lists of explanatory variables.

Explain carefully about what is the marginal product of each of these methods. Old-fashioned OLS and logit give some predictions – are these other methods better overall or in particular cases? Do they tend to make the same sort of errors or different ones? (If different then perhaps you can create an ensemble of models?)


